Bayesian Nonparametric System Reliability\\ using Sets of Priors

### 2016-01-12 ijar-npb-ms.tex
minor typos and tweaks

Let us denote the random failure time of component number $i$ of type $k$ by $T^k_i$, $i = 1, \ldots m_k$.
The failure time distribution can be written in terms of the cdf $F^k(t) = P(T^k_i \le t)$,
or in terms of the reliability function $R^k(t) = P(T^k > t) = 1 - F^k(t)$,
also known as the survival function.
For a nonparametric description of $R^k(t)$,
we consider a set of time points $t$, $t \in {\cal T} = \{t_1, \ldots, \tmax\}$.

@@

For discussing this reparametrisation, we will drop the super- and subscript $k$ and $t$ for a while.
Instead of $\alpha$ and $\beta$, we consider the parameters $\nz$ and $\yz$, where
\begin{align}
\nz &= \alpha + \beta &
&\text{and} &
\yz &= \frac{\alpha}{\alpha + \beta} \,,
\label{eq:reparam}
\end{align}
or vice versa, $\alpha = \nz\yz$ and $\beta = \nz(1-\yz)$.
The upper index ${}\uz$ is used to identify $\nz$ and $\yz$ as prior parameter values,
in contrast to their posterior values $\nn$ and $\yn$
obtained after observing $n$ failure times, see below.
$\nz$ and $\yz$ are sometimes called \emph{canonical} parameters,
identified from rewriting the density in a canonical form;
see, e.g., \citet[pp.~202 and 272f]{2000:bernardosmith}, or \citet[\S 1.2.3.1]{2013:diss-gw}.
This canonical form gives a joint structure to all conjugate priors in exponential families.

From the properties of the Beta distribution,
it follows that $\yz = \E[p]$ is the prior expectation for the functioning probability $p$,
and that the higher $\nz$, the more probability weight will be concentrated around $\yz$,
as $\V(p) = \frac{\yz (1-\yz)}{\nz + 1}$.
Furthermore, $\nz$ can be interpreted as a pseudocount or prior strength,
this is clear from the interpretation of $\alpha$ and $\beta$ and \eqref{eq:reparam},
and will be further illustrated below through revisiting the posterior
%based on observed lifetimes $\vec{t} = (t_1, \ldots, t_{n})$.
given $s$ of $n$ components functioning.
We have that
$p \mid s$ is Beta distributed with the updated parameters
\begin{align}
\nn &= \nz + n\,, &
\yn &= \frac{\nz}{\nz + n} \cdot \yz + \frac{n}{\nz + n} \cdot \frac{s}{n}\,.
\label{eq:nyupdate}
\end{align}
After seeing that $s$ out of $n$ components function (at time $t$),
the posterior mean $\yn$ for $p$ is a weighted average of
the prior mean $\yz$ and $s/n$ (the fraction of functioning components in the data),
with the weights $\nz$ and $n$, respectively.
We see that $\nz$ plays the same role for the prior mean $\yz$
as the sample size $n$ for the observed mean $s/n$,
thus the notion of pseudocount.
Indeed, the higher $\nz$, the higher the weight for $\yz$
in the weighted average calculation of $\yn$,
so $\nz$ gives the strength of the prior as compared to the sample size $n$.
%The parameters $\nz$ and $\yz$ have a more intuitive interprtation,

@@

In absence of prior-data conflict, 
$\PktN$ shrinks in the $\ykt$ dimension;
how much it shrinks depends on $\nktz \in [\nktzl, \nktzu]$,
leading to the so-called spotlight shape depicted in Figure~\ref{fig:paramsets} (left).
As $\yktn$ gives the posterior expectation for the functioning probability $p_t^k$,
shorter $\yktn$ intervals mean more precise knowledge about $p_t^k$.
Also, the variance interval for $p_t^k$ will decrease
as the Beta distributions in $\MktN$ will be more pointed
due to the increase of $\nktz$ to $\nktn$.

@@

Based on these shapes and \eqref{eq:nyupdate}, it is possible to deduce the following expressions
for the lower and upper bounds of $\yktn$:
\begin{align}
\begin{aligned}
\min_{\PktN} \yktn &=
 \begin{cases}
 %\dfrac{\nktzu \yktzl + s^k_t}{\nktzu + n_k} & \text{if } s^k_t/n_k \ge \yktzl \\
 \big(\nktzu \yktzl + s^k_t\big) / \big(\nktzu + n_k\big) & \text{if } s^k_t/n_k \ge \yktzl \\
 %\dfrac{\nktzl \yktzl + s^k_t}{\nktzl + n_k} & \text{if } s^k_t/n_k <   \yktzl
 \big(\nktzl \yktzl + s^k_t\big) / \big(\nktzl + n_k\big) & \text{if } s^k_t/n_k <   \yktzl
 \end{cases}\,,\\
\max_{\PktN} \yktn &=
 \begin{cases}
 %\dfrac{\nktzu \yktzu + s^k_t}{\nktzu + n_k} & \text{if } s^k_t/n_k \le \yktzu \\
 \big(\nktzu \yktzu + s^k_t\big) / \big(\nktzu + n_k\big) & \text{if } s^k_t/n_k \le \yktzu \\
 %\dfrac{\nktzl \yktzu + s^k_t}{\nktzl + n_k} & \text{if } s^k_t/n_k >   \yktzu
 \big(\nktzl \yktzu + s^k_t\big) / \big(\nktzl + n_k\big) & \text{if } s^k_t/n_k >   \yktzu
 \end{cases}\,.
\end{aligned}
\label{eq:ysetupdate}
\end{align}
The lower bound for $\yktn$ is always attained for $\yktzl$, the upper bound for $\yktzu$.
Note that when $s^k_t/n_k \in [\yktzl, \yktzu]$,
both the lower and the upper bound for $\yktn$ are attained using $\nktzu$,
this corresponds to the spotlight shape.
When $s^k_t/n_k \not\in [\yktzl, \yktzu]$,
due to the banana shape
we will get one of the bounds for $\yktn$ using $\nktzl$.

>>>>

Let us denote the random failure time of component number $i$ of type $k$ by $T^k_i$, $i = 1, \ldots, m_k$.
The failure time distribution can be written in terms of the cdf $F^k(t) = P(T^k_i \le t)$,
or in terms of the reliability function $R^k(t) = P(T^k_i > t) = 1 - F^k(t)$,
also known as the survival function.
For a nonparametric description of $R^k(t)$,
we consider a set of time points $t$, $t \in {\cal T} = \{t_1, \ldots, \tmax\}$.

@@

For discussing this reparametrisation, we will drop the super- and subscript $k$ and $t$ for a while.
Instead of $\alpha$ and $\beta$, we consider the parameters $\nz$ and $\yz$, where
\begin{align}
\nz &= \alpha + \beta &
&\text{and} &
\yz &= \frac{\alpha}{\alpha + \beta} \,,
\label{eq:reparam}
\end{align}
or vice versa, $\alpha = \nz\yz$ and $\beta = \nz(1-\yz)$.
The upper index ${}\uz$ is used to identify $\nz$ and $\yz$ as prior parameter values,
in contrast to their posterior values $\nn$ and $\yn$
obtained after observing $n$ failure times, see below.
$\nz$ and $\yz$ are sometimes called \emph{canonical} parameters,
identified from rewriting the density in a canonical form;
see, e.g., \cite[pp.~202 and 272f]{2000:bernardosmith}, or \cite[\S 1.2.3.1]{2013:diss-gw}.
This canonical form gives a common structure to all conjugate priors in exponential families.

From the properties of the Beta distribution,
it follows that $\yz = \E[p]$ is the prior expectation for the functioning probability $p$,
and that the higher $\nz$, the more probability weight will be concentrated around $\yz$,
as $\V(p) = \frac{\yz (1-\yz)}{\nz + 1}$.
Furthermore, $\nz$ can be interpreted as a pseudocount or prior strength,
this is clear from the interpretation of $\alpha$ and $\beta$ and \eqref{eq:reparam},
and will be further illustrated below through revisiting the posterior
%based on observed lifetimes $\vec{t} = (t_1, \ldots, t_{n})$.
given that $s$ of $n$ components function.
We have that
$p \mid s$ is Beta distributed with the updated parameters
\begin{align}
\nn &= \nz + n\,, &
\yn &= \frac{\nz}{\nz + n} \cdot \yz + \frac{n}{\nz + n} \cdot \frac{s}{n}\,.
\label{eq:nyupdate}
\end{align}
After seeing that $s$ out of $n$ components function (at time $t$),
the posterior mean $\yn$ for $p$ is a weighted average of
the prior mean $\yz$ and $s/n$ (the fraction of functioning components in the data),
with the weights $\nz$ and $n$, respectively.
We see that $\nz$ plays the same role for the prior mean $\yz$
as the sample size $n$ for the observed mean $s/n$,
thus the notion of pseudocount.
Indeed, the higher $\nz$, the higher the weight for $\yz$
in the weighted average calculation of $\yn$,
so $\nz$ gives the strength of the prior as compared to the sample size $n$.
%The parameters $\nz$ and $\yz$ have a more intuitive interprtation,

@@

In absence of prior-data conflict, 
$\PktN$ shrinks in the $\ykt$ dimension;
how much it shrinks depends on $\nktz \in [\nktzl, \nktzu]$,
leading to the so-called spotlight shape depicted in Figure~\ref{fig:paramsets} (left).
As $\yktn$ gives the posterior expectation for the functioning probability $p_t^k$,
shorter $\yktn$ intervals mean more precise knowledge about $p_t^k$.
Also, the variance interval for $p_t^k$ will shorten and shift towards zero,
as the Beta distributions in $\MktN$ will be more pointed
due to the increase of $\nktz$ to $\nktn$.

@@

Based on these shapes and \eqref{eq:nyupdate}, it is possible to deduce the following expressions
for the lower and upper bounds of $\yktn$:
\begin{align}
\begin{aligned}
\min_{\PktN} \yktn &=
 \begin{cases}
 %\dfrac{\nktzu \yktzl + s^k_t}{\nktzu + n_k} & \text{if } s^k_t/n_k \ge \yktzl \\
 \big(\nktzu \yktzl + s^k_t\big) / \big(\nktzu + n_k\big) & \text{if } s^k_t/n_k \ge \yktzl \\
 %\dfrac{\nktzl \yktzl + s^k_t}{\nktzl + n_k} & \text{if } s^k_t/n_k <   \yktzl
 \big(\nktzl \yktzl + s^k_t\big) / \big(\nktzl + n_k\big) & \text{if } s^k_t/n_k <   \yktzl
 \end{cases}\,,\\
\max_{\PktN} \yktn &=
 \begin{cases}
 %\dfrac{\nktzu \yktzu + s^k_t}{\nktzu + n_k} & \text{if } s^k_t/n_k \le \yktzu \\
 \big(\nktzu \yktzu + s^k_t\big) / \big(\nktzu + n_k\big) & \text{if } s^k_t/n_k \le \yktzu \\
 %\dfrac{\nktzl \yktzu + s^k_t}{\nktzl + n_k} & \text{if } s^k_t/n_k >   \yktzu
 \big(\nktzl \yktzu + s^k_t\big) / \big(\nktzl + n_k\big) & \text{if } s^k_t/n_k >   \yktzu
 \end{cases}\,.
\end{aligned}
\label{eq:ysetupdate}
\end{align}
The lower bound for $\yktn$ is always attained for $\yktzl$, the upper bound for $\yktzu$.
Note that when $s^k_t/n_k \in [\yktzl, \yktzu]$,
both the lower and the upper bound for $\yktn$ are attained using $\nktzu$,
this corresponds to the spotlight shape.
When $s^k_t/n_k \not\in [\yktzl, \yktzu]$,
the banana shape indicates that
we will get one of the bounds for $\yktn$ using $\nktzl$.
